Increasing User Engagement in a Volunteer-Based Ephemeral Evolutionary Computation System

### 2017-02-17 2017-EUROPAR/Gamification/paper.tex
Experimental setup

Gamification techniques in a volunteer context seeks to persuade 
users to use their natural desire to compete, learn and socialize in 
given non-game context application \cite{deterding2011game, hamari2014does}.  
Some works give a form of reward to users, these include 
points \cite{sutter2010browse}, achievement badges or levels \cite{hamari2011framework}, 
the filling of a progress bar \cite{o2010get}, or providing the user with virtual currency.
By Making the rewards for  tasks achievements visible to other players or 
providing leader boards are ways of encouraging players to compete \cite{hickman2010total}. 
Competitio can also have problematic consequences, which can result in
negative conduct, low cooperation and collaboration, or disadvantaging certain player demographics
such as women \cite{kumar2013gamification}. Another techniques to gamification 
is to make existing tasks feel more like games \cite{deterding2010just}. 
Some techniques used in this approach include adding meaningful choice, 
onboarding with a tutorial, increasing challenge, and adding narrative \cite{mcgonigal2011reality}.

\section{Experiment}
\label{sec:experiments}
In the browser, each page visiting the experiment loads an HTML Worker
that runs a local island of an evolutionary algorithm to solve a
multimodal problem called {\em l-trap}, which has been used extensively 
as a benchmark for evolutionary algorithms \cite{fernandes2009using,nijssen2003analysis}. 
This function counts the number of bits in a sequence $l$ and assigns
the local maximum $a$ if it has got 0 bits and the global maximum $b$ if it has $l$
bits. As seen in Figure \ref{fig:trap} the fitness falls into a {\em trap} 
as the number of bits is increased, decreasing linearly until a change in slope 
is reached at point $z$, adding deceptive component for evolutionary algorithms. 
To increase the difficulty trap functions can be concatenated.
In our case we have used $40$ concatenated traps. The trap function  defined as:   
\[ f(u)= 
    \begin{cases} 
      \frac{a}{z}(z-u) & \text{if } u\leq z\\
      \frac{b}{l-z} (u-z)& \text{otherwise} 
   \end{cases}
\]
\begin{figure*}[t]
    \centering
        \includegraphics[width=3in]{img/trap.png}
    \caption{Trap function.
    }
    \label{fig:trap}
\end{figure*}

Each local GA had the following parameters, the initial population was randomly generated 
with sizes between 128 and 256 individuals, the period to send individuals to the server
was set at 100 local generations, the parameters for the trap function are $l$ = 4 and $b$ = 2.

>>>>

Gamification techniques in a volunteer context seeks to persuade 
users to use their natural desire to compete, learn and socialize in 
given non-game context application \cite{deterding2011game,hamari2014does}.  
Some works give a form of reward to users, these include 
points \cite{sutter2010browse}, achievement badges or levels \cite{hamari2011framework}, 
the filling of a progress bar \cite{o2010get}, or providing the user with virtual currency.
By Making the rewards for  tasks achievements visible to other players or 
providing leader boards are ways of encouraging players to compete \cite{hickman2010total}. 
Competition can also have problematic consequences, which can result in
negative conduct, low cooperation and collaboration, or disadvantaging certain player demographics
such as women \cite{kumar2013gamification}. Another techniques to gamification 
is to make existing tasks feel more like games \cite{deterding2010just}. 
Some techniques used in this approach include adding meaningful choice, 
on-boarding with a tutorial, increasing challenge, and adding narrative \cite{mcgonigal2011reality}.

\section{Experiment}
\label{sec:experiments}

In this experiment a gamification technique is applied to 
the {\sf NodIO} volunteer computing framework, using the 
particular version described in \cite{DBLP:conf/gecco/MereloCGCRV16,2016arXiv160101607M}.
The objective of the experiment is to test the kind of impact 
applying a gamification technique has on user engagement in this
particular application. In order to apply a rewarding system
user authentication had to be developed first. In earlier designs
this functionality was not desired because it was seen as a barrier 
for participation. After all the advantage of a browser based volunteer system
is precisely the minimum amount of user intervention needed to start.
The last thing a user wants to see is yet another registration form.
Then, the first design decision for this version is that registration is optional
and not promoted by the interface. There is only a button with a ``Login/Register'' label,
using the graphical interface shown in Figure \ref{fig:login}. 
The registration form is very basic, having only the user-name and 
password (with out confirmation) as the required fields. Email is optional
and is not verified. The problem of users that forgot their password
happened, and it was resolved by sending an email to the administrators
or some users just created another account.

The gamification technique employed in this work is based on a rewarding mechanism  
\cite{dubois2013understanding}. In general rewards  consist of a reputation system 
with score points, levels and leader boards. Points are awarded to users in response of 
the accomplishment of certain activities that need to be encouraged. In these case
a point is awarded for each PUT sent to the server. Levels are a long
term achievement, in this case the level depends on the score:
\[ \text{level}(\text{score};a,b)= 
    \begin{cases} 
      0,                                    &  \text{score}\leq a\\    
      2(\frac{\text{score}-a}{b-a})^{2},    &  a\leq \text{score}\leq \frac{a+b}{2}\\
      1-2(\frac{\text{score}-b}{b-a})^{2},  & \frac{a+b}{2} \leq \text{score}\leq b\\
      1,                                    & \text{score}\geq b 
   \end{cases}
\]
The function returns a normalized value that is multiplied by the maximum level 
in this case 100. The variables $a$=100 and $b$=7360 give the slope of the function as
seen in Figure~\ref{fig:s}. The function was set to give users a rapid increase of 
levels at the beginning.
\begin{figure*}[t]
    \centering
        \includegraphics[width=3in]{img/s.png}
    \caption{$S$ function used to assign the user's level.
    }
    \label{fig:s}
\end{figure*}

After login, registered users are welcomed by their name, and they can see
their score and level. All users can see in a modal dialog the leader board.
While the modal is opened, the scores are refreshed every second. All other 
functionality is available to all users and is the same as the previous 
{\sf NodIO} version, showing the state of the current algorithm. 
In the front page a link to the open source code of the experiment was also 
available \url{https://github.com/lucero21/login-master}

\subsection{Problem}
In the browser, each page visiting the experiment loads an HTML Worker
that runs a local island of an evolutionary algorithm to solve a
multi-modal problem called {\em l-trap}, which has been used extensively 
as a benchmark for evolutionary algorithms \cite{fernandes2009using,nijssen2003analysis}. 
This function counts the number of bits in a sequence $l$ and assigns
the local maximum $a$ if it has got 0 bits and the global maximum $b$ if it has $l$
bits. As seen in Figure \ref{fig:trap} the fitness falls into a {\em trap} 
as the number of bits is increased, decreasing linearly until a change in slope 
is reached at point $z$, adding deceptive component for evolutionary algorithms. 
To increase the difficulty trap functions can be concatenated.
In our case we have used $40$ concatenated traps. The trap function  defined as:   
\[ f(u)= 
    \begin{cases} 
      \frac{a}{z}(z-u) & \text{if } u\leq z\\
      \frac{b}{l-z} (u-z)& \text{otherwise} 
   \end{cases}
\]
\begin{figure*}[t]
    \centering
        \includegraphics[width=3in]{img/trap.png}
    \caption{Trap function.
    }
    \label{fig:trap}
\end{figure*}

Each local GA had the following parameters, the initial population was randomly generated 
with sizes between 128 and 256 individuals, the period to send individuals to the server
was set at 100 local generations, the parameters for the trap function are $l$ = 4,
$a$=1, $b$ = 2, $z$=3 and a chromosome length of 160 bits.

In order to begin the experiment a call to participation was issued on November 23th, 
2016 through the authors social networks:
``Asking again for your help, we are conducting a computational experiment
that requires computer power. Can we borrow some of your CPU? just visit the web page (link)
and leave the tab open. Be part of the TOP TEN, register so we can track your participation.
The experiment will run until November 27th, Thanks!''. In the message, the registration
to the application is presented, and also a deadline of the experiment. When the
experiment was over, users visiting the page were presented with a thank you message 
and a static leader board.

\begin{figure*}[t]
    \centering
        \includegraphics[width=5in]{img/login.png}
    \caption{Comparison of 30 runs of the 128 Bit OneMax problem. 
    Box-plot of the number of evaluations needed for solution, with a 2, 6 and 12 workers
    homogeneous configuration on the left side, and Heterogeneous configuration on the
    right side of each.
    }
    \label{fig:login}
\end{figure*}
